A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-05-06 2022/Sensors/kNN.tex
second comment

@@

Note that the centre  $y^c(x)$ of the $M$-th partition subcube containing point $y(x)$ can be regarded as an approximation to $y(x)$ with the accuracy of $\varepsilon = 2^{-(M+1)}$, i.e.
\begin{equation}
\max_{1 \leq j \leq N} \left| y_j^c(x)-y_j(x) \right| \leq \varepsilon = 2^{-(M+1)}.
\end{equation}

The function $y^c(x)$ maps a uniform grid with step $2^{-NM}$, constructed in the interval $[0,1]$, onto a uniform grid with step $2^{-M}$, constructed in the hypercube $D$. A constructive way of establishing such a correspondence is described and theoretically justified in \cite{Strongin2000, Sergeyev2013}.

@@

Consider a family of Peano curves
\begin{equation}
Y_L(x)=\{ y^0(x), y^1(x), ..., y^L(x) \}
\end{equation} 
instead of a single curve $y(x)$. We will construct the family of curves as follows. Let us introduce a family of hypercubes
\begin{equation}\label{hypercubes}
D_i = \{ y \in R^N: -2^{-1} \leq y_i + 2^{-l} \leq 3 \cdot 2^{-1},  1 \leq i \leq N \}, \; 0 \leq l \leq L,
\end{equation} 
where the hypercube $D_{l+1}$ is obtained by shifting the hypercube $D_l$ along the main diagonal by step $2^{-l}$ for each coordinate.

@@

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Figure~\ref{fig2}), then, by introducing an additional function 
\[
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\] 
the original hypercube $D$ can be represented as
\[
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\] 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\[
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\]

@@

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
\[
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\] 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
\[
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\]

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Investigation of neighbour search times on real-world data}

>>>>

%Баркалов
Отметим, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}. Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.

@@

Note that the centre  $y^c(x)$ of the $M$-th partition subcube containing point $y(x)$ can be regarded as an approximation to $y(x)$ with the accuracy of $\varepsilon = 2^{-(M+1)}$, i.e.
\begin{equation}\label{prec}
\max_{1 \leq j \leq N} \left| y_j^c(x)-y_j(x) \right| \leq \varepsilon = 2^{-(M+1)}.
\end{equation}

The function $y^c(x)$ maps a uniform grid with step $2^{-NM}$, constructed in the interval $[0,1]$, onto a uniform grid with step $2^{-M}$, constructed in the hypercube $D$. A constructive way of establishing such a correspondence is described and theoretically justified in \cite{Strongin2000, Sergeyev2013}. 
%Баркалов
Алгоритмическая сложность вычисления значения функции $y^c(x)$ зависит от размерности $N$, от уровня разбиения $M$ и составляется величину $O(M\cdot N)$.

@@

Consider a family of Peano curves
\begin{equation}\label{EvSet}
Y_L(x)=\{ y^0(x), y^1(x), ..., y^L(x) \}
\end{equation} 
instead of a single curve $y(x)$. We will construct the family of curves as follows. Let us introduce a family of hypercubes
\begin{equation}\label{hypercubes}
D_i = \{ y \in R^N: -2^{-1} \leq y_i + 2^{-l} \leq 3 \cdot 2^{-1},  1 \leq i \leq N \}, \; 0 \leq l \leq L,
\end{equation} 
where the hypercube $D_{l+1}$ is obtained by shifting the hypercube $D_l$ along the main diagonal by step $2^{-l}$ for each coordinate.

@@

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Figure~\ref{fig2}), then, by introducing an additional function 
\[
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\] 
the original hypercube $D$ can be represented as
\[
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\] 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\[
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\] 
%Баркалов
Алгоритмическая сложность вычисления всех прообразов $x^l \in [0, 1], \; 0 \leq l \leq L,$ зависит от размерности $N$, от уровня разбиения $M$, числа отображений $L$ и составляется величину $O(M\cdot N \cdot L)$.

@@

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
\begin{equation}\label{y_prime}
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\end{equation} 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
\[
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\]

@@

%Баркалов
Как было сказано выше, вместо теоретической Hilbert curve $y(x)$ на практике используется evolvent $y^c(x)$, которая является приближением к $y(x)$ с точностью порядка $2^{-(M+1)}$, определяемой соотношением (\ref{prec}). При этом параметр $M$ может выбираться для каждой решаемой задачи индивидуально, исходя из точности представления исходных данных.
Например, данные в  DS-Skin dataset (Section \ref{RealData}) представляют собой координаты пикселей изображения с разрешением 256 на 256. При отображении данных на стандартный гиперкуб $D$ from (\ref{D}) расстояние по координате между двумя точками не может быть меньше $1/256$. Использование параметра $M=10$ для построения evolvent будет достаточным в этом случае. 
В случае же присвоения численных значений небольшому числу categorical features (как это было сделано для CarEvaluation dataset, see Section \ref{RealData}) расстояние по координате между объектами будет большое, значение параметра $M$ можно снизить (в экспериментах использовалось значение $M=7$).

Максимальное число одновременно используемых отображений в множестве $Y_L(x)$ from (\ref{EvSet}) также будет определяться точностью исходных данных, от которой зависит точность построения разверток $y^i(x)$.
Будем счистать, что каждое из отображений $y^i(x)$ построено с использованием одного и того же значения $M$. Тогда значение $L$ должно удовлетворять неравенству $L \leq M$. В противном случае точки $y' = y^l(x'),\; y''=y^l(x'')$ из условия (\ref{y_prime}) будут соответствовать центру одного и того же подкуба при $p>M$, т.е. будет выполняться равенство $y' = y''$  что приведет к нарушению условий Теоремы 1.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Investigation of neighbour search times on real-world data}\label{RealData}
